
# Activity Monitoring Study
#### Author: "Aditya"
#### Date: "June 4, 2017"


#### Summary: *This Report captures the Analysis done to analyse and interpret the data collected from personal activity monitoring device of an Individual for 2 months.*

## The below R code chunk will load the R Packages

```{r loading packages,echo=TRUE, message=FALSE, warning=FALSE}
        library("markdown")
        library("rmarkdown")
        library("knitr")
        library("dplyr")
        library("ggplot2")
```

 
## Setting R Markdown Global options
```{r setoptions}
        opts_chunk$set(echo=TRUE, results = "asis", message=FALSE, error=FALSE, warning=FALSE, cache = TRUE)
```

  
## Downloading the Source Data and Reading into R

```{r Importing Data}

        file_url <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
        setwd <- "G:/Data Science Project/Reproducible Research/Week 2/Assignment"
        download.file(url=file_url,destfile = "Activity.zip",method = "libcurl")
        unzip("Activity.zip")
        activity<- read.csv(file ="activity.csv",header = TRUE, stringsAsFactors = FALSE)

```

## The below R code helps in understanding the structure & summary Statistics of activity table. 

```{r}
# Dimensions of the Activity Table
        dim(activity)

# Structure of the Activity Table
        str(activity)
        
# Summary statistics of the Activity Table
        summary(activity)
```

## Identifying Missing Values in the Entire Table                
```{r}
        sum(is.na(activity))
        mean(is.na(activity))
```

## Identifying Missing Values Column/ Variable wise                
```{r}        
        colSums(is.na(activity))
        colMeans(is.na(activity))
```

## Transforming format of Date variable 
```{r}        
        #activity$date=as.POSIXlt(activity$date)
    
        activity$date=as.Date(activity$date)
```

## Histogram of the total number of steps taken each day

```{r Total Steps per Day1}
        activity_no_na <- activity[!rowSums(is.na(activity)),]
    
        total_steps <- aggregate(steps~date,data=activity_no_na, FUN = sum,na.rm = FALSE)

ggplot(total_steps, aes(steps))+ geom_histogram(binwidth = 5000) + labs(x="Steps per Day",y="Frequency", title="Histogram of Total no of Steps per Day") + theme(plot.title = element_text(hjust=0.5))
 
        
```

## Mean and median number of steps taken each day

```{r Mean & Median of steps per day}        
        round(mean(total_steps$steps))
        median(total_steps$steps)
```

## Average No of Steps by Interval
```{r}
    avg_steps <- aggregate(steps~interval, activity_no_na, FUN=mean)

ggplot(avg_steps, aes(x=interval,y=steps))+geom_line(col="red")+labs(x="5- minute Inteval", y="Average Step Count", title="Average No of Steps by Interval") + theme(plot.title = element_text(hjust = 0.5))
        
```

## The 5-minute interval on average contains the maximum number of steps
```{r Max no of steps interval}      
    avg_steps<- arrange(avg_steps, desc(steps))
    paste("The 5- minute interval -- ", avg_steps[1,"interval"], ", has maximum no of steps i.e. " ,round(avg_steps[1,"steps"])   )

```

## Imputing missing data -- Using the Average steps of each 5- minute interval

```{r Imputing Missing Data}      

    activity_new <- merge.data.frame(activity, avg_steps, by.x = "interval", by.y="interval")
    for (i in 1: nrow(activity_new))
    {   
        activity_new$steps[i]<- if (is.na(activity_new$steps.x[i])) {activity_new$steps.y[i]}                                   else {activity_new$steps.x[i]}
    }
    activity_new1 <- select(activity_new,c(interval,date,steps))

```


## Histogram of the total number of steps taken each day after missing values are imputed
```{r Total Steps per Day Imputed values}

        total_steps1 <- aggregate(steps~date,data=activity_new1, FUN = sum)

ggplot(total_steps1, aes(steps))+ geom_histogram(binwidth = 5000) + labs(x="Steps per Day",y="Frequency", title="Histogram of Total no of Steps per Day") + theme(plot.title = element_text(hjust=0.5))

```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r panel plot}
    activity_new1$date<- as.POSIXlt(activity_new1$date)
    activity_new2<- activity_new1
    x <- activity_new2$date$wday
    for(i in 1:length(x))
        {activity_new2$day_flag[i] <- if(x[i] %in% c(6,7)) {"weekend"} else {"weekday"}}

    avg_steps2 <- aggregate(steps~interval+day_flag, activity_new2, FUN = sum)
    
    ggplot(avg_steps2, aes(x=interval, y=steps)) + geom_line(col="red") + facet_wrap(~day_flag) +labs(x="5- minute Inteval", y="Average Step Count", title="Average No of Steps by Interval across Weekday and Weekend") + theme(plot.title = element_text(hjust = 0.5))
    
```


